---
title: "pCO2 temperature conversion"
author: "Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r packages}

library(tidyverse)
library(seacarb)
library(patchwork)



```




# Temperature coefficient

## Takahasi estimate

As a first step, I reanalyse the temperature coefficient derived by Takahashi et al (1993), based on the values in table A1.

```{r read_Takahashi_data, fig.asp=1.5}

df <- read_csv(here::here("data", "Takahashi_1993_Table_A1.csv"))

model <- lm(ln_pCO2 ~ tem, data = df)
model
summary(model)


df <- df %>% 
  mutate(ln_pCO2_rec = log(pCO2),
         d_ln_pCO2_rec = ln_pCO2 - ln_pCO2_rec,
         ln_pCO2_model = predict(model, df),
         d_ln_pCO2_model = ln_pCO2 - ln_pCO2_model,
         pCO2_model = exp(ln_pCO2_model),
         d_pCO2 = pCO2 - pCO2_model)

p_fit <- df %>% 
  ggplot(aes(tem, ln_pCO2))+
  stat_smooth(method = "lm", col="red", se=FALSE)+
  geom_point()+
  labs(x="Temperature °C", y=expression(ln~(pCO[2])))

p_residual <- df %>% 
  ggplot(aes(tem, d_ln_pCO2_model))+
  geom_hline(yintercept = 0)+
  geom_point()+
  labs(x="Temperature °C", y=expression(Delta~ln~(pCO[2])[obs-model]))

p_fit / p_residual


```

The linear model fit gives a slope of `r round(model$coefficients[2],4)` which is slightly lower compared to the slope determined by the authors. The standard deviation of the model residuals amounts to `r round(sd(model$residuals),4)`, which is equal to `r round(100*sd(model$residuals)/0.0423,1)` % of the determined slope.

```{r remove_objects}

rm(df, model, p_fit, p_residual)

```


## Dissociation constants

As an alterantive approach, the temperature coefficient was computed based on the CO~2~ dissociation constants in seawater.Computation were done with the R package `seacarb`. Computations were performed for a range of absolute temperature and pCO~2~ levels. Alkalinity and salintiy were assumed constant (AT = 2300 µmol/kg; S = 35). For each set of conditions, following calculations were performed:

1. DIC was calculated from AT and pCO~2,equi~
2. Temperature was reduced by 1K
3. pCO~2,insitu~ was calculated for reduced temperature, based on AT and DIC
4. dln(pCO2)/dT was computed

Computations were performed for different sets of CO~2~ dissociation constant k1 and k2 available in seacarb:

- "l" Lueker et al. (2000)
- "m10" Millero (2010)
- "w14" Waters et al. (2014)
- "r" Roy et al. (1993)


```{r compute_T_coeff, fig.asp=1.3}

df <- expand_grid(AT   = 2300,
                  sal    = 35,
                  Tequ  = seq(0,30,5),
                  pCO2equ = seq(200,600,100),
                  k1k2 = c("l", "m10", "w14", "r"))


df <- df %>% 
  mutate(AT = AT*1e-6,
         DIC = carb(flag=24, var1=pCO2equ, var2=AT, S=sal, T=Tequ, k1k2=k1k2)[,16],
         SST = Tequ - 1,
         pCO2_insitu = carb(flag=15, var1=AT, var2=DIC, S=sal, T=SST, k1k2=k1k2)[,9],
         d_pCO2 = pCO2equ - pCO2_insitu,
         T_coeff = (log(pCO2equ) - log(pCO2_insitu))/(Tequ-SST),
         d_T_coeff_rel = 100*T_coeff/0.0423)


df %>% 
  ggplot(aes(Tequ, T_coeff, col=as.factor(pCO2equ)))+
  geom_hline(yintercept = 0.0423, col="red")+
  annotate("text", x = 25, y = 0.043, label = "0.0423", col="red")+
  geom_line()+
  geom_point()+
  scale_color_viridis_d(name = expression(pCO[2]~(µatm)))+
  labs(x=expression(T[equi]~("°C")), y=expression(dln(pCO[2])/dT))+
  facet_wrap(~k1k2, labeller = label_both)+
  theme(legend.position = "bottom")

df %>% 
  ggplot(aes(Tequ, d_T_coeff_rel, col=as.factor(pCO2equ)))+
  geom_hline(yintercept = 100, col="red")+
  geom_line()+
  geom_point()+
  scale_color_viridis_d(name = expression(pCO[2]~(µatm)))+
  labs(x=expression(T[equi]~("°C")), y=expression(dln(pCO[2])/dT~relativ~to~0.0423))+
  facet_wrap(~k1k2, labeller = label_both)+
  theme(legend.position = "bottom")

```

## Model fit

A model was fitted to the temperature coefficient determined with the constants by Lueker et al. (2000), as a function of temperature and pCO~2~. The chosen model accounts for a quadratic dependence of the temperature coeffcient on temperature and a linear dependence on pCO~2~


```{r fit_T_coeff, fig.asp=1}

df <- df %>% 
  filter(k1k2 == "l") %>% 
  select(Tequ, pCO2equ, T_coeff)


#library(mgcv)
#model <- gam(T_coeff ~ s(Tequ) + s(pCO2equ), data = df)

model <- lm(T_coeff ~ poly(Tequ,2) + pCO2equ, data = df)

model

model %>% 
  saveRDS(here::here("/data/model","T_coeff_model.rds"))


df <- df %>% 
  mutate(T_coeff_model = predict(model),
         d_T_coeff = T_coeff - T_coeff_model)

model_plot <- expand_grid(Tequ  = seq(0,30,0.1),
                  pCO2equ = seq(200,600,100))

model_plot <- model_plot %>% 
  mutate(T_coeff = predict(model, model_plot))


p1 <- df %>% 
  ggplot(aes(Tequ, T_coeff, fill=as.factor(pCO2equ)))+
  geom_hline(yintercept = 0.0423, col="red")+
  annotate("text", x = 25, y = 0.043, label = "0.0423", col="red")+
  geom_line(data = model_plot, (aes(Tequ, T_coeff, col=as.factor(pCO2equ))))+
  geom_point(shape=21)+
  scale_color_viridis_d(name = expression(pCO[2]~(µatm)))+
  scale_fill_viridis_d(name = expression(pCO[2]~(µatm)))+
  scale_y_continuous(breaks = seq(0, 1,0.001), sec.axis = sec_axis(~ . *100/0.0423, name = "Relative to 0.0423 (%)"))+
  labs(x=expression(T[equi]~("°C")), y=expression(dln(pCO[2])/dT))


p2 <- df %>% 
  ggplot(aes(Tequ, d_T_coeff, fill=as.factor(pCO2equ)))+
  geom_hline(yintercept = 0)+
  geom_point(shape=21)+
  scale_fill_viridis_d(name = expression(pCO[2]~(µatm)))+
  labs(x=expression(T[equi]~("°C")), y=expression(Delta~(dln(pCO[2])/dT)[calc - model]))+
  theme(legend.position = "none")


(p1 / p2) + 
  plot_layout(guides = 'collect')

```


### Comparison to Schneider

The model fitted here was compared to the one fitted by Bernd Schneider, based on the CO~2~ dissociation constants by Mojica Prieto and Millero (2002), which are currently not available in `seacarb`.


```{r compare_fit_T_coeff_Schneider}

df <- df %>% 
  mutate(T_coeff_Schneider = 3.92*1e-6*Tequ^2 - 4.27*1e-4*Tequ - 4.384*1e-6*pCO2equ + 4.814*1e-2,
         d_T_coeff_Schneider = T_coeff_Schneider - T_coeff_model)

model_plot <- expand_grid(Tequ  = seq(0,30,0.1),
                  pCO2equ = seq(200,600,100))

model_plot <- model_plot %>% 
  mutate(T_coeff_Schneider = 3.92*1e-6*Tequ^2 - 4.27*1e-4*Tequ - 4.384*1e-6*pCO2equ + 4.814*1e-2,)


df %>% 
  ggplot(aes(Tequ, T_coeff, fill=as.factor(pCO2equ)))+
  geom_hline(yintercept = 0.0423, col="red")+
  annotate("text", x = 25, y = 0.043, label = "0.0423", col="red")+
  geom_line(data = model_plot, (aes(Tequ, T_coeff_Schneider,
                                    col=as.factor(pCO2equ),
                                    linetype = "Schneider")))+
  scale_linetype(name="model")+
  scale_shape_manual(name="model", values = 21)+
  geom_point(aes(shape="Mueller/Luecker"))+
  scale_color_viridis_d(name = expression(pCO[2]~(µatm)))+
  scale_fill_viridis_d(name = expression(pCO[2]~(µatm)))+
  scale_y_continuous(breaks = seq(0, 1,0.001))+
  labs(x=expression(T[equi]~("°C")), y=expression(dln(pCO[2])/dT))


```



