---
title: "Revision of temperature conversion within SOCAT"
author: "Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r packages}

library(tidyverse)
library(metR)
library(patchwork)

# theme_set(theme(
#   panel.background = element_rect(
#     colour = "black")))

```


# SOCAT data base

## Read and subset data

This study is based on SOCATv2019.tsv, thus excluding flag E data stoyellow in a seperate file.
Limitation to observations with given equilibrator temperature probably restricts observations to those made on VOS (excluding bouys).

```{r read_subset_SOCAT_data, eval=FALSE}

# identify the number of rows to skip by playing around with the import data set functionality in RStudio

df <- read_tsv(here::here("/data", "SOCATv2019.tsv"), skip = 5859)

df <- df %>% 
  select(4:12, 14:16, 19, 21:26)

df <- df %>% 
  rename(lon = `longitude [dec.deg.E]`, 
         lat = `latitude [dec.deg.N]`,
         SST = `SST [deg.C]`, 
         Tequ = `Tequ [deg.C]`, 
         dep = `ETOPO2_depth [m]`,
         dist_land = `dist_to_land [km]`,
         GVCO2 = `GVCO2 [umol/mol]`,
         fCO2rec = `fCO2rec [uatm]`)

nrow_total <- nrow(df)

df <- df %>% 
  filter(!is.na(Tequ))

nrow_total_Tequ <- nrow(df)

df <- df %>% 
  filter(!is.na(SST))

nrow_total_Tequ_SST <- nrow(df)

df <- df %>% 
  filter(fCO2rec < 700, fCO2rec > 100)

nrow_total_Tequ_SST_fCO2rec <- nrow(df)

  df_sub_5e6 <- df %>% 
    sample_n(size = 5e6, replace = FALSE)
  
  df_sub_1e6 <- df %>% 
    sample_n(size = 1e6, replace = FALSE)
  
  df_sub_1e5 <- df %>% 
    sample_n(size = 1e5, replace = FALSE)
  
  df_sub_1e4 <- df %>% 
    sample_n(size = 1e4, replace = FALSE)
  
  df  %>% write_csv(here::here("/data/_summarized_data", "SOCATv2019_Tequi_not_NaN.csv"))
  df_sub_5e6  %>% write_csv(here::here("/data/_summarized_data", "SOCATv2019_random_5e6.csv"))
  df_sub_1e6  %>% write_csv(here::here("/data/_summarized_data", "SOCATv2019_random_1e6.csv"))
  df_sub_1e5  %>% write_csv(here::here("/data/_summarized_data", "SOCATv2019_random_1e5.csv"))
  df_sub_1e4  %>% write_csv(here::here("/data/_summarized_data", "SOCATv2019_random_1e4.csv"))


```

The following analysis is based on the subsetted SOCAT data base and 100 µatm < fCO2 < 700 µatm. 

```{r read_modified_SOCAT_data}


#df <- read_csv(here::here("/data/_summarized_data", "SOCATv2019_Tequi_not_NaN.csv"))
df <- read_csv(here::here("/data/_summarized_data", "SOCATv2019_random_5e6.csv"))

#floor_decade    = function(value){ return(value - value %% 10) }


df %>%
  filter(fCO2rec < 800) %>%
  ggplot(aes(fCO2rec))+
  geom_vline(xintercept = c(100, 700), col="red")+
  geom_histogram()

df <- df %>%
  mutate(d_Tem = SST-Tequ,
         lon = if_else(lon > 190, lon-360, lon))



df <- df %>% 
  select(3, 8, 9, 11, 12, 17, 20)


```

## Global maps


```{r plot_SOCAT_maps_all}

maps_SOCAT <- df %>% 
  mutate(lat_int = cut(lat,
                       breaks = seq(-90,90,1),
                       labels = seq(-89.5,89.5,1)),
         lat_int = as.numeric(as.character(lat_int)),
         lon_int = cut(lon,
                       breaks = seq(-170,190,1),
                       labels = seq(-169.5,189.5,1)),
         lon_int = as.numeric(as.character(lon_int))) %>% 
  select(-c(lat, lon))

maps_SOCAT_all <- maps_SOCAT%>% 
  select(-mon) %>% 
  group_by(lat_int, lon_int) %>% 
  summarise_all("mean")


mapWorld <- borders("world", colour="gray60", fill="gray60") # create a layer of borders

maps_SOCAT_all %>% 
  ggplot(aes(lon_int, lat_int, fill=SST))+
  mapWorld+
  geom_tile()+
  scale_fill_viridis_c(name = "SST (°C)", option = "plasma", na.value = "black")+
  labs(x="Lon (°E)", y="Lat (°N)")+
  scale_x_continuous(breaks = seq(-180, 180, 30), limits = c(-170, 190.5))+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  coord_quickmap(expand = FALSE)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "map_SST.png"), width = 9, height = 5)


# maps_SOCAT_all %>% 
#   filter(fCO2rec<500) %>% 
#   ggplot(aes(fCO2rec))+
#   geom_histogram()

maps_SOCAT_all %>% 
  ggplot(aes(lon_int, lat_int, fill=fCO2rec))+
  mapWorld+
  geom_tile()+
  scale_fill_viridis_c(name = expression(fCO[2]~(µatm)), limits = c(200, 500), na.value = "black")+
  labs(x="Lon (°E)", y="Lat (°N)")+
  scale_x_continuous(breaks = seq(-180, 180, 30), limits = c(-170, 190.5))+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  coord_quickmap(expand = FALSE)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "map_pCO2.png"), width = 9, height = 5)

maps_SOCAT_all %>% 
  ggplot(aes(lon_int, lat_int, fill=d_Tem))+
  mapWorld+
  geom_tile()+
  scale_fill_divergent(name = expression(Delta~T[SST-equi]~(K)), na.value = "black")+
  labs(x="Lon (°E)", y="Lat (°N)")+
  scale_x_continuous(breaks = seq(-180, 180, 30), limits = c(-170, 190.5))+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  coord_quickmap(expand = FALSE)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "map_d_Tem.png"), width = 9, height = 5)

#rm(maps_SOCAT, maps_SOCAT_all, mapWorld)

```


```{r plot_SOCAT_maps_monthly, eval=FALSE}

maps_SOCAT_monthly <- maps_SOCAT %>% 
  group_by(lat_int, lon_int, mon) %>% 
  summarise_all("mean")

maps_SOCAT_monthly %>% 
  ggplot(aes(lon_int, lat_int, fill=d_Tem))+
  mapWorld+
  geom_tile()+
  scale_fill_viridis_c(name = "SST (°C)")+
  labs(x="Lon (°E)", y="Lat (°N)")+
  scale_x_continuous(breaks = seq(-180, 180, 30), limits = c(-170, 190.5))+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  coord_quickmap(expand = FALSE)+
  facet_wrap(~mon, ncol = 3, labeller = label_both)


```

## Latitudinal patterns

```{r plot_SOCAT_lat_patterns}

lat_pattern <- df %>% 
  mutate(lat = cut(lat,
                       breaks = seq(-90,90,5),
                       labels = seq(-87.5,87.5,5)),
         lat = as.numeric(as.character(lat)),
         SST = cut(SST,
                       breaks = seq(-4,40,1),
                       labels = seq(-3.5,39.5,1)),
         SST = as.numeric(as.character(SST)),
         fCO2rec = cut(fCO2rec,
                       breaks = seq(100,700,20),
                       labels = seq(110,690,20)),
         fCO2rec = as.numeric(as.character(fCO2rec)),
         d_Tem = cut(d_Tem,
                       breaks = seq(-5,5,0.2),
                       labels = seq(-4.9,4.9,0.2)),
         d_Tem = as.numeric(as.character(d_Tem))) %>% 
  #drop_na() %>% 
  select(-c(mon, lon, Tequ))

lat_mean <- df %>% 
  mutate(lat = cut(lat,
                       breaks = seq(-90,90,5),
                       labels = seq(-87.5,87.5,5)),
         lat = as.numeric(as.character(lat)))


SST_lat_pattern <- lat_pattern %>% 
  group_by(lat, SST) %>% 
  summarise(nr=n()) %>% 
  ungroup()

SST_lat_mean <- lat_mean %>% 
  group_by(lat) %>% 
  summarise(SST = mean(SST, na.rm = TRUE)) %>% 
  ungroup()


SST_lat_pattern %>% 
  ggplot()+
  geom_raster(aes(SST, lat, fill=log10(nr)))+
  geom_path(data = SST_lat_mean, aes(SST, lat, col="Mean"), size=1.5)+
  geom_hline(yintercept = 0)+
  scale_fill_continuous(name = "lg (nr obs)")+
  scale_color_manual(name = "", values = "yellow")+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  labs(y="Lat (°N)", x="SST (°C)")+
  coord_cartesian(expand = 0)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "lat_SST.png"), width = 5, height = 5)


fCO2rec_lat_pattern <- lat_pattern %>% 
  group_by(lat, fCO2rec) %>% 
  summarise(nr=n()) %>% 
  ungroup()

fCO2rec_lat_mean <- lat_mean %>% 
  group_by(lat) %>% 
  summarise(fCO2rec = mean(fCO2rec, na.rm = TRUE)) %>% 
  ungroup()

fCO2rec_lat_pattern %>% 
  ggplot()+
  geom_raster(aes(fCO2rec, lat, fill=log10(nr)))+
  geom_path(data = fCO2rec_lat_mean, aes(fCO2rec, lat, col="Mean"), size=1.5)+
  geom_hline(yintercept = 0)+
  scale_fill_continuous(name = "lg (nr obs)")+
  scale_color_manual(name = "", values = "yellow")+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  labs(y="Lat (°N)", x=expression(fCO[2]~(µatm)))+
  coord_cartesian(expand = 0)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "lat_pCO2.png"), width = 5, height = 5)


d_Tem_lat_pattern <- lat_pattern %>% 
  group_by(lat, d_Tem) %>% 
  summarise(nr=n()) %>% 
  ungroup()

d_Tem_lat_mean <- lat_mean %>% 
  group_by(lat) %>% 
  summarise(d_Tem = mean(d_Tem, na.rm = TRUE)) %>% 
  ungroup()

d_Tem_lat_pattern %>% 
  ggplot()+
  geom_raster(aes(d_Tem, lat, fill=log10(nr)))+
  geom_path(data = d_Tem_lat_mean, aes(d_Tem, lat, col="Mean"), size=1.5)+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  scale_fill_continuous(name = "lg (nr obs)")+
  scale_color_manual(name = "", values = "yellow")+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  labs(y="Lat (°N)", x=expression(Delta~T[SST-equi]~(K)))+
  coord_cartesian(expand = 0)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "lat_d_Tem.png"), width = 5, height = 5)

rm(lat_pattern, d_Tem_lat_pattern, fCO2rec_lat_pattern, SST_lat_pattern)

```

```{r fCO2_vs_d_Tem, eval=FALSE}

d_Tem_pCO2_mean <- df %>% 
   mutate(d_Tem = cut(d_Tem,
                       breaks = seq(-5,5,0.2),
                       labels = seq(-4.9,4.9,0.2)),
          d_Tem = as.numeric(as.character(d_Tem))) %>% 
  group_by(d_Tem) %>% 
  summarise(fCO2rec = mean(fCO2rec, na.rm = TRUE)) %>% 
  ungroup()

df %>% 
  ggplot()+
  geom_point(aes(d_Tem, fCO2rec))+
  geom_vline(xintercept = 0)+
  ylim(200,500)

```




# Temperature conversion

fCO~2~ in the equilibrator is backcalculated from fCO2rec based on the Takahashi temperature dependence of pCO~2~.
fCO2rec is than re-calculated with the temperature- and pCO~2~ dependent conversion factor determined before.

```{r recalc_fCO2}

model <- 
  readRDS(here::here("/data/model","T_coeff_model.rds"))



df <- df %>%
  mutate(pCO2equ = fCO2rec)

df <- df %>%
  mutate(fCO2equ = fCO2rec / exp(0.0423*d_Tem),
         t_coeff = predict(model, df),
         fCO2rec_Mueller = fCO2equ * exp(t_coeff*d_Tem),
         d_fCO2 = fCO2rec_Mueller - fCO2rec) %>% 
  select(-pCO2equ)


```


```{r maps_of_SOCAT_recalc}

maps_SOCAT_recalc <- df %>% 
  mutate(lat_int = cut(lat,
                       breaks = seq(-90,90,1),
                       labels = seq(-89.5,89.5,1)),
         lat_int = as.numeric(as.character(lat_int)),
         lon_int = cut(lon,
                       breaks = seq(-170,190,1),
                       labels = seq(-169.5,189.5,1)),
         lon_int = as.numeric(as.character(lon_int))) %>% 
  select(lat_int, lon_int, t_coeff, d_fCO2)

maps_SOCAT_all <- maps_SOCAT_recalc %>% 
  group_by(lat_int, lon_int) %>% 
  summarise_all("mean")


mapWorld <- borders("world", colour="gray60", fill="gray60") # create a layer of borders


maps_SOCAT_all %>% 
  ggplot(aes(lon_int, lat_int, fill=100*t_coeff/0.0423))+
  mapWorld+
  geom_tile()+
  scale_fill_divergent(name = expression(dln(pCO[2])/dT~("%"~of~0.0423)), midpoint = 100, na.value = "black")+
  labs(x="Lon (°E)", y="Lat (°N)")+
  scale_x_continuous(breaks = seq(-180, 180, 30), limits = c(-170, 190.5))+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  coord_quickmap(expand = FALSE)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "map_d_T_coeff.png"), width = 9, height = 5)

maps_SOCAT_all %>% 
  filter(d_fCO2 > -2.5, d_fCO2 <2.5) %>% 
  ggplot(aes(lon_int, lat_int, fill=d_fCO2))+
  mapWorld+
  geom_tile()+
  scale_fill_divergent(name = expression(Delta~pCO["2,Müller-SOCAT"]~(µatm)), na.value = "black")+
  labs(x="Lon (°E)", y="Lat (°N)")+
  scale_x_continuous(breaks = seq(-180, 180, 30), limits = c(-170, 190.5))+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  coord_quickmap(expand = FALSE)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "map_d_fCO2.png"), width = 9, height = 5)


```

```{r plot_SOCAT_recalc_lat_patterns}

lat_pattern <- df %>% 
  mutate(lat = cut(lat,
                       breaks = seq(-90,90,5),
                       labels = seq(-87.5,87.5,5)),
         lat = as.numeric(as.character(lat)),
         d_fCO2 = cut(d_fCO2,
                       breaks = seq(-5,5,0.2),
                       labels = seq(-4.9,4.9,0.2)),
         d_fCO2 = as.numeric(as.character(d_fCO2)),
         t_coeff = cut(100*t_coeff/0.0423,
                       breaks = seq(50,150,1),
                       labels = seq(50.5,149.5,1)),
         t_coeff = as.numeric(as.character(t_coeff)))

lat_mean <- df %>% 
  mutate(lat = cut(lat,
                       breaks = seq(-90,90,5),
                       labels = seq(-87.5,87.5,5)),
         lat = as.numeric(as.character(lat)))


t_coeff_lat_pattern <- lat_pattern %>% 
  group_by(lat, t_coeff) %>% 
  summarise(nr=n()) %>% 
  ungroup()

t_coeff_lat_mean <- lat_mean %>% 
  group_by(lat) %>% 
  summarise(t_coeff = mean(100*t_coeff/0.0423, na.rm = TRUE)) %>% 
  ungroup()

t_coeff_lat_pattern %>% 
  ggplot()+
  geom_raster(aes(t_coeff, lat, fill=log10(nr)))+
  geom_path(data = t_coeff_lat_mean, aes(t_coeff, lat, col="Mean"), size=1.5)+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 100)+
  scale_fill_continuous(name = "lg (nr obs)")+
  scale_color_manual(name = "", values = "yellow")+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  labs(y="Lat (°N)", x=expression(dln(pCO[2])/dT~("%"~of~0.0423)))+
  coord_cartesian(expand = 0)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "lat_t_coeff.png"), width = 5, height = 5)



d_fCO2_lat_pattern <- lat_pattern %>% 
  group_by(lat, d_fCO2) %>% 
  summarise(nr=n()) %>% 
  ungroup()

d_fCO2_lat_mean <- lat_mean %>% 
  group_by(lat) %>% 
  summarise(d_fCO2 = mean(d_fCO2, na.rm = TRUE)) %>% 
  ungroup()

d_fCO2_lat_pattern %>% 
  ggplot()+
  geom_raster(aes(d_fCO2, lat, fill=log10(nr)))+
  geom_path(data = d_fCO2_lat_mean, aes(d_fCO2, lat, col="Mean"), size=1.5)+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  scale_fill_continuous(name = "lg (nr obs)")+
  scale_color_manual(name = "", values = "yellow")+
  scale_y_continuous(breaks = seq(-90, 90, 30), limits = c(-80, 90))+
  scale_x_continuous(limits = c(-2.5, 2.5))+
  labs(y="Lat (°N)", x=expression(Delta~pCO["2,Müller-SOCAT"]~(µatm)))+
  coord_cartesian(expand = 0)+
  theme(legend.position = "bottom")

ggsave(here::here("output/figures", "lat_d_fCO2.png"), width = 5, height = 5)


```


# Open Tasks

- Check if rows with equilibrator temperature NaN refer to bouy data
- why are there SST NaN rows?
